 \documentclass[UKenglish]{ifimaster}  %% ... or USenglish or norsk
\usepackage[utf8]{inputenc}           %% ... or latin1
\usepackage[T1]{fontenc,url}
\urlstyle{sf}
\usepackage{babel,textcomp,csquotes,duomasterforside,varioref,graphicx}
\usepackage[backend=biber,style=numeric-comp, sorting=none]{biblatex}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{textgreek}
\usepackage{float}
\usepackage{graphicx}
\usepackage{chemformula}
\usepackage{cleveref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{chemformula}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
\usepackage{siunitx}

\raggedbottom %%reduces the gaps between the paragraphs
\title{An ab initio study of the band gaps of high-entropy silicides based on \ch{FeSi2}}        %% ... or whatever
%\subtitle{}         %% ... if any
\author{Jørn-Marcus Høylo-Rosenberg}                      %% ... or whoever 

\let\oldAA\AA
\renewcommand{\AA}{\text{\normalfont\oldAA}}

\addbibresource{bibliography.bib}  
\graphicspath{{figures/}}
\DeclareUnicodeCharacter{2212}{-}

\includeonly
{
   	chapter/chapter_2,
    chapter/chapter_3,
    chapter/chapter_4,
    chapter/chapter_5,
    chapter/chapter_6,
    chapter/chapter_7,
    chapter/chapter_8,
    chapter/chapter_11
}


\begin{document}
\duoforside[dept={Department of Chemistry},   %% ... or your department
  program={Materials Science for Energy and Nanotechnology},  %% ... or your programme
  long]                                        %% ... or long

\frontmatter{}  

\chapter*{Abstract}                   %% ... or Sammendrag or Samandrag

In this work we have set out to examine the band gaps of potential high-entropy stabilized transition metal silicides, based on the \ch{FeSi2} semiconductor. Here, we have performed a first principles study utilizing density functional theory in combination with the special quasi-random structures method. The band gaps has been evaluated primarily with the PBE GGA functional, but also the SCAN meta-GGA functional and hybrid functional HSE06 has been applied in this project. Potential alloys has been examined by generating a set of distinct supercells of the \ch{FeSi2} structure, where the iron sites are populated quasi-randomly by a combination of 3d transition metals. The alloys we have studied in this project are comprised of combinations of Cr, Fe, Mn, Ni, Co and Ti, most emphasis has been put on the Cr-Fe-Mn-Ni-Si system. In the \ch{(CrFeMnNi)Si2} composition, we can report several semiconducting supercells with band gaps ranging between 0 - 0.05 eV using PBE GGA. The band gaps displayed significant spin polarization, in the spin up direction most supercells pointed to a gap around 0.3 eV. Contrary, the band gaps in spin down varied between 0 - 0.05 eV. Accordingly, we found a finite magnetic moment in this alloy equal to 0.083 $\mu_B$, this was attributed mainly to chromium and manganese atoms in the lattice. Successive simulations of alloys based on the Cr-Fe-Mn-Ni-Si system with non-equimolar distribution of 3d elements, resulted most frequently in half-metallic structures with a spin up band gap ranging between 0.1 eV and 0.5 eV. Of the different compositions we tested, the ones either rich in manganese and/or poor in chromium showed the most promise with respect to the band gap. In particular, we report a PBE total band gap equal to 0.1 eV in the \ch{Cr3Fe5Mn5Ni3Si32} composition. Lastly, we looked at compositions where either Cr, Mn or Ni were replaced by either cobalt or titanium. This yielded predominately metallic compounds, where the lack of band gaps could be ascribed to defect states at the band edges. 


\chapter*{Acknowledgments}                    %% ... or Forord
I would like to express my gratitude towards my supervisor Ole Martin Løvvik. For always being available and prioritize my work from the first day to the last. Coming into this project, I had no prior knowledge of density functional theory or quantum mechanical modeling. So, for your guidance in this project I am truly thankful, and know that this project would have been impossible without you. I also feel as if I have made a friend and someone I could talk through throughout this project. Due to the oddity of these recent years with the covid-19 outbreak and self-isolation this have been very important to my sanity and mental health combined with the difficulty involved in writing a master theses. On this same note, I would like to thank all my friends and family members that have been there and believed in me, and always reminding me to stay positive. 

\tableofcontents{}
\listoffigures{}
\listoftables{}

\mainmatter{}

\chapter{Introduction}                  %% ... or Bakgrunn

A major concern in the energy sector is the loss of energy as waste heat. For instance, \cite{thermo} estimated that about 60 percent of the energy produced in USA in 2012 was lost as waste heat. Contrary to common energy sources that suffer from this problem of losing energy to heat, thermoelectrcity can generate electricity from heat by what's known as the Seebeck effect \cite{wikipedia_thermoelectric_2020}. Hence, thermoelecticity can have many promising applications, for example one could imagine a conjunction between thermoelectric generators and mechanical generators to recover a portion of the waste heat produced by the latter. However, currently thermoelectricity is of limited practical use due to a lack of efficient thermoelectric materials.  This can be explained from the thermoelectric figure of merit, composed of the Seebeck Coefficient $S$, electrical conductivity $\sigma$, working temperature $T$ and the thermal conductivity $\kappa$, in the following relation \cite{bandgap2Dirac} 

\begin{equation}
zT = \frac{S^2\sigma T}{\kappa}. 
\end{equation}

An efficient thermoelectric material consists of three dependent variables: A high Seebeck coefficient, high electrical conductivity, and low thermal conductivity. A popular strategy to achieve this relation is the phonon-glass electron crystal approach \cite{SnyderG.Jeffrey2008Ctm}, in which one will introduce complexity and anharmonicity to the crystal lattice by some means to increase levels of phonon scattering, without also increasing the scattering of electrons. Furthermore, good thermoelectrics are narrow gap semiconductors. \cite{TE_gap} found highest zT values for semiconductors with band gap between 6-10 $k_B$T, which at room temperature equals around 0.2 eV.   

In this project we will look at high-entropy alloys as potential thermoelectric materials. High-entropy alloys are a novel material class that extends the concept of traditional binary-alloys such as steel, to multi-component alloys. The name "high-entropy" stems from that the materials are stabilized by the high configurational entropy caused by alloying multiple constituents. These materials are in particular known for their strong mechanical properties \cite{hea}, such as high strength at elevated temperatures. Moreover, evidence of low thermal conductivity originating from the disordered crystal structure. Hence, provided that we can find narrow gap high-entropy alloys, these could be relevant and promising high zT thermoelectric materials.   

Specifically, in this project we will look at high-entropy silicides. These are a lesser studied subclass of high-entropy alloys. The first experimentally synthesized high-entropy silicide was found in 2019 by Gild. et al \cite{GILD2019337}. He studied a first of its kind single-phase high-entropy disilicide \ch{(MoNbTaW)Si2}. This silicide adopted the hexagonal C40 crystal structure, and displayed low thermal conductivity compared to conventional disilicides in the equivalent crystal structure. Additionally, very recently a master student at the University of Oslo managed to synthesize three phases of non-cubic high-entropy silicides based on Si, Co, Cr, Fe, and Ni in both hexagonal and orthorhombic symmetries \cite{mari}. We base this project on silicides because firstly silicon and various transition metal-silicides are environmentally sound and silicon in particular is heavily applied in micro-electronic devices and renewable energy sources such as solar power, thus silicon based alloys could readily be implemented into these technologies. Secondly, transition metal silicides offer a good range of initial compounds with suitable band gaps for thermoelectrics \cite{silicde}.   

This project is performed by ab initio methods in the framework of density functional theory. The disordered structure of high-entropy alloys are handled with a numerical method called special quasi-random structures. The particular materials investigated in this project will be based on the semiconductor $\beta$-\ch{FeSi2}. We will mainly investigate the band gap of such alloys, for this we employ three classes of functionals in DFT: GGA, meta-GGA and hybrid functionals. In addition, we will perform a brief analysis of the stability and magnetism. The potential high-entropy silicides are studied by  constructing five distinct supercells of the master compound, where the iron sites are populated by a quasi-random distribution between 3d transition metals. Special emphasis are placed on alloys consisting of Cr, Fe, Mn and Ni, as these along with silicon are sustainable non-toxic elements.  

We begin this project by reviewing the fundamentals and properties of high-entropy alloys, thereafter we present a theoretical description of the computational methods used in this project, SQS, and DFT. Next, we discuss various practical aspects of DFT, such as the exchange-correlation functionals and numerical convergence. Following, is a section on the computational settings and dependencies necessary to reproduce the results presented in this thesis. Finally, we present and discuss the results, which begins with the high-entropy silicide \ch{(CrFeMnNi)Si2}, followed by various derivatives and alternatives to this system.  In the end we give a brief conclusion of our work and discuss possible directions for future research.


\part{Theory}                    %% ... or ??
\include{chapter/chapter_2}
\include{chapter/chapter_3}
\include{chapter/chapter_4}

\part{Method}
\include{chapter/chapter_5}
\include{chapter/chapter_6}
 
\part{Results and Discussion}
\include{chapter/chapter_7}
\include{chapter/chapter_8}


\part{Conclusion and future work}
\chapter{Conclusion}

High-entropy alloys has emerged as a rapidly growing interest in materials science from the large range of possibilities and flexibility. In this project we have set out to study the possibility of a narrow gap semiconducting high-entropy silicide with top of the line computational power and methods, in the prospect of discovering next-generation efficient thermoelectric materials. The materials in mind were based on the \ch{FeSi2} semiconductor, and modeled with the SQS method in the framework of DFT. To study the potential semiconductors we generated five distinct configurations of each composition and investigated the band gap, mainly with PBE GGA.  

Our most successful effort was based around a 48-atom model of chromium, iron, manganese, and nickel in the Cr:Fe:Mn:Ni:Si system. In the equimolar composition \ch{(CrFeMnNi)Si2}, we observed large variation of the band gap between the five distinct configurations. In the spin up direction, the band gap ranged from 0.08 - 0.34 eV. The most stable supercell was a half-metal with a spin up band gap of 0.34 eV, while the other four supercells displayed small spin down gaps between 0.03 eV and 0.05 eV as well. Calculations with SCAN and HSE06 produced unpredictable outputs that varied greatly between SQSs, and contained a larger degree of uncertainty compared to PBE. However, with HSE06 we found spin up band gaps of 0.55 eV and 0.71 eV in two respective supercells, and a total band gap of 0.18 eV in another. With the SCAN functional we found total band gaps around 0.1 in three SQSs, one metallic supercell, and one half-metallic structure with a spin down band gap of 0.1 eV.   

Following the equimolar composition, we conducted a brief exploration of the quaternary phase diagram of this composition. Contrary to the equimolar system, we found here more frequently half-metallic structures. By analyzing the projected density of states we were able to relate the metallic spin down channel to a dominant presence of manganese states around the Fermi energy.  Inside the number of different compositions that we tested, we found evidence for that there exists a positive relation between the band gap, most notably in spin up, and compositions poor in chromium and/or rich in manganese. Specifically, we located two promising compositions: \ch{Cr3Fe3Mn7Ni3Si32} and \ch{Cr3Fe5Mn5Ni3Si32}, where the most stable supercell of the former displayed a spin up band gap of 0.47 eV, and the latter a total band gap of 0.1 eV, with PBE.

Lastly, we tested compositions of the \ch{FeSi2} based alloys consisting of cobalt/titanium in place of either Cr, Mn or Ni. Based on the most stable configurations of these compositions, the only band gaps we found in this experiment was a spin up band gap of 0.46 eV in \ch{(CrFeMnCo)Si2}. By manually investigating the calculated eigenvalues of the different compositions, we were able to relate the metallic compounds to defect states. This is a familiar term in random alloys, in which states at the band edges deviate slightly from completely full/empty occupancy.  

In accordance with the spin polarization of the band gap, we noted a finite magnetic moment in most alloys.  Distinctly, the equimolar alloy contained a magnetic moment of $0.083 \mu_B$ in all five configurations of the equimolar alloy. Upon investigation of the local magnetic moments, we discovered that the magnetism was mainly attributed to chromium and manganese atoms in the lattice. This was found as a general trend for all alloys based on the \ch{FeSi2} structure. From this severe spin polarization of the band gap, a possible application of the Cr-Fe-Mn-Ni-Si alloys, could be as spintronics \cite{spintronic}. 

In terms of thermoelectric applications, we find limited success in respect to the band gap. Recalling that good thermoelectrics are semiconductors with band gaps around 0.2 eV, the band gaps of the \ch{(CrFeMnNi)Si2} alloy are too small. However, we have found instances more suitable, such as the HSE06 band gap of 0.18 eV in one of the supercells, and the 0.1 eV band gap in \ch{Cr3Fe5Mn5Ni3Si32}. Furthermore, we can expect that the band gap of the real material to be larger than the PBE value, thus the Cr-Fe-Mn-Ni-Si system could have potential application as a thermoelectric.

\chapter{Future work}

In future studies, I would first and foremost try to resolve several of the uncertainties circling the results presented in this project. This would include a more thoughtful investigation of the true ground state of the \ch{(CrFeMnNi)Si2} alloy. For example conduct a test of multiple configurations and a larger number of SQS sizes, in order reduce variance and obtain a converged value of the band gap. Secondly, to decide on the most representative configuration, we should have put more effort into correctly/optimally specify the magnetic configuration of each SQS. Also, a possible extension could be to study excited states, to account for entropic contributions to the stability.  Moreover, a test of different crystal structures in addition to the orthorombic CMCE space group should be done, as the local minima method of DFT does not guarantee that this is the most stable conformation of the alloy.

The overarching motivation of this thesis has been to simply locate a band gap in high-entropy silicides. Now that we have located such a compound, in future work we could devote more effort into specifically analyze the band gap. Potentially try to plot and analyze the band structure of the alloys and qualitatively scrutinize the band gap by different band gap specific functionals. Furthermore, we performed only a very limited search of the quaternary phase diagram of the Cr-Fe-Mn-Ni-Si system, but found promising results in the direction of less chromium and more manganese. Thus, a future study on more deliberate compositions in these directions would also be interesting. Provided that these results yielded positive outcomes of the band gap, suitable for thermoelectric application, future work could focus on a complete study of this material as a thermoelectric, and consider properties such as the Seebeck coefficient, electrical conductivity, and thermal conductivity. 

\appendix
\include{chapter/chapter_11}

\backmatter{}
\printbibliography

\end{document}
